---
title: "1.meta_aim2"
author: "Emily Yeo"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
pacman::p_load(knitr, data.table, dplyr, tidyr,readxl,
               readr, car, LongituRF)
```

## Load Data
```{r}
m1_dir <- "/Users/emily/projects/research/Stanislawski/comps/mutli-omic-predictions/data/clinical/transformed/aim2"
test <- read.csv(paste(m1_dir, "a2_test_samples_standard_clinical.csv", sep = "/"))
train <- read.csv(paste(m1_dir, "a2_train_samples_standard_clinical.csv", sep = "/"))
```

# Long RF {.tabset}

## Demo long RF
my_data is a list of what you'd expect:
Y (response vector), 
X (matrix of fixed effects predictors), 
Z (matrix of random-effects predictors), 
id (vector of sample identifier), 
time (vector of time measurements). 
To fit random forest model simply

```{r}
my_data <- DataLongGenerator(n=50,p=6,G=6)
model <- REEMforest(X=my_data$X,
                    Y=my_data$Y,
                    Z=my_data$Z,
                    time=my_data$time,
                    id=my_data$id,
                    sto="BM",mtry=2)
```

## Making the Z matrix with real data 
Simulate longitudinal data according to the semi-parametric stochastic mixed-effects model given. 

When the function DataLongGenerator() creates Z, it's a random uniform data in a matrix. The actual coding is
```{r}
Z <- as.matrix(cbind(rep(1, length(f)), 2 * runif(length(f))))
```

Where f represents the length of the matrices that represent each of the elements. 
In your example, you used 2 groups of 97 participants with 12 fixed effects.

```{r}
data_train <- DataLongGenerator(n=97, p=12) # Generate the data
# Let's see the output :
w <- which(data$id==1)
plot(data$time[w],data$Y[w],
     type="l",
     ylim=c(min(data$Y),
            max(data$Y)), 
     col="grey")
for (i in unique(data$id)){
w <- which(data$id==i)
lines(data$time[w],data$Y[w], col='grey')
}
# Let's see the fixed effects predictors:
par(mfrow=c(2,3), mar=c(2,3,3,2))
for (i in 1:ncol(data$X)){
w <- which(data$id==1)
plot(data$time[w],data$X[w,i], col="grey",ylim=c(min(data$X[,i]),
max(data$X[,i])),xlim=c(1,max(data$time)),main=latex2exp::TeX(paste0("$X^{(",i,")}$")))
for (k in unique(data$id)){
w <- which(data$id==k)
lines(data$time[w],data$X[w,i], col="grey")
}
}
par(oldpar)
options(oldopt)
```

# {-}



